---
title: "MLC Data Processing"
author: "Emily Bovee"
date: "1/30/2019"
output: html_document
---
Note to self: The primary difference between this script and the 1.24.19 script is that I reararnged the order of the deidentification.
The purpose of this is to be able to troubleshoot the merge issue without the issue of compromising student privacy.

Overview of Workflow
(1) Import 10 tabs from Excel multi-tab document
(2) Add a variable to designate semester (tabs are separated by semester)
(3) Add study IDs
(4) Deidentify
(5) Remove people who are not part of my study

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##########################################
#Setting working directory

  #check what working directory currently is
getwd()

  #set working directory to point R in the file where the data is
setwd("/Volumes/Linnen Lab/MSU Engineering/PROJECTS/Emily Dissertation")

  #check to make sure setting the working directory above worked.
getwd()

##########################################
#Loading packages for analysis

library(tidyverse)
library(readxl)
library(tidylog)
```
##Overview
  This document includes the processing of the MLC data. In its raw form, this dataset was a multi-tab spreadsheet with information from each semester contained in its own tab.
  
  My goal in processing it is to have all the data in one unified dataframe.
  
  The thing that will make this easier than processing the Center data is that the variable names are the same across tabs, so once I get everything in R, I can process it in the same way.
  
  Another step that I will take is to remove the people who are not part of my study.

## 1. Reading in Files
```{r}

##1. Read MLC data into R

  FS15 <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 1)
  SS16 <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 2)
  US16A <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 3)
  US16B <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 4)
  FS16 <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 5)
  SS17 <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 6)
  US17A <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 7)
  US17B <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 8)
  FS17 <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 9)
  SS18 <- read_excel("MLC_data_Emily_Bovee.xlsx", sheet = 10)
 
studyIDs <- read_excel("Dissertation Master File - 12.12.18.xlsx")
```    
## 2. Add semester variable to tag which semester each is for

```{r}
#Append a SEMESTER variable to EACH of the above files to tag which semester it is
  #so, for FS15 the variable would say FS15, etc.
  
FS15 <- FS15 %>%
  mutate(semester = "FS15")

SS16 <- SS16 %>%
  mutate(semester = "SS16")

US16A <- US16A %>%
  mutate(semester = "US16A")

US16B <- US16B %>%
  mutate(semester = "US16B")

FS16 <- FS16 %>%
  mutate(semester = "FS16")

SS17 <- SS17 %>%
  mutate(semester = "SS17")

US17A <- US17A %>%
  mutate(semester = "US17A")

US17B <- US17B %>%
  mutate(semester = "US17B")

FS17 <- FS17 %>%
  mutate(semester = "FS17")

SS18 <- SS18 %>%
  mutate(semester = "SS18")

  #check the dataframes created by the above syntax to make sure that each semester variables is correct

```

## 3. Add Study IDs
```{r}
#I already loaded the file at the top of the code

#rename the "PID" variable to match the file I'll be joining, which calls it "Pid"
#rename the other variables so there are no spaces

studyIDs <- studyIDs %>% 
  rename(Pid = PID,
         study_id = `Study ID`,
         net_id = `Net ID`,
         first = `First name`,
         last = `Last name`)

#merging study IDs to the 10 component files

FS15 <- left_join(FS15, studyIDs, by = "Pid")

SS16 <- left_join(SS16, studyIDs, by = "Pid")

US16A <- left_join(US16A, studyIDs, by = "Pid")

US16B <- left_join(US16B, studyIDs, by = "Pid")

FS16 <- left_join(FS16, studyIDs, by = "Pid")

SS17 <- left_join(SS17, studyIDs, by = "Pid")

US17A <- left_join(US17A, studyIDs, by = "Pid")

US17B <- left_join(US17B, studyIDs, by = "Pid")

FS17 <- left_join(FS17, studyIDs, by = "Pid")

SS18 <- left_join(SS18, studyIDs, by = "Pid")
```


## 4. Deidentify - remove PIDs, Net IDs, first and last name, and cohort
```{r}
#Remove all identifying variables
drop <- c("Pid", "net_id", "first", "last", "COHORT")

FS15 <- FS15  %>% 
  select(-one_of(drop))

SS16 <- SS16  %>% 
  select(-one_of(drop))

US16A <- US16A  %>% 
  select(-one_of(drop))

US16B <- US16B  %>% 
  select(-one_of(drop))

FS16 <- FS16  %>% 
  select(-one_of(drop))

SS17 <- SS17  %>% 
  select(-one_of(drop))

US17A <- US17A  %>% 
  select(-one_of(drop))

US17B <- US17B  %>% 
  select(-one_of(drop))

FS17 <- FS17  %>% 
  select(-one_of(drop))

SS18 <- SS18 %>% 
  select(-one_of(drop))
```

## 5.  Remove people who are not part of my study
```{r}
  #delete if they don't have a study ID

FS15 <- FS15 %>% 
  filter(study_id != "") 

SS16 <- SS16  %>% 
  filter(study_id != "") 

US16A <- US16A  %>% 
  filter(study_id != "") 

US16B <- US16B  %>% 
  filter(study_id != "") 

FS16 <- FS16  %>% 
  filter(study_id != "") 

SS17 <- SS17  %>% 
  filter(study_id != "") 

US17A <- US17A  %>% 
  filter(study_id != "") 

US17B <- US17B  %>% 
  filter(study_id != "") 

FS17 <- FS17  %>% 
  filter(study_id != "") 

SS18 <- SS18 %>% 
  filter(study_id != "") 
```

# Remove people who were after the Spring survey started (i.e. after 3/22/18)

```{r}
SS18 <- SS18 %>% 
  filter(Date < as.Date("2018-03-22"))

#str(SS18$Date)

#For filtering by date
# https://blog.exploratory.io/filter-with-date-function-ce8e84be680
```

# 6. Vertical Merge: get all observations in one dataframe with rbind
```{r}

#Check dimensions of everything before merge
  #note the number of rows

dim (FS15)
  # 3099

dim (FS16)
  # 4348

dim (FS17)
  # 1198

dim (SS16)
  # 4336

dim (SS17)
  # 3987

dim (SS18)
  # 294

dim (US16A)
  # 78

dim (US16B)
  # 71

dim (US17A)
  # 67

dim (US17B)
  # 31

#Expected Count of rows in new dataset (sum row dimensions of all above)
31 + 67 + 71 + 78 + 294 + 3987 + 4336 + 1198 + 4348 + 3099
  #17509

MLC <- rbind(FS15, FS16, FS17, SS16, SS17, SS18, US16A, US16B, US17A, US17B)

#FURTHER READING 

#More about For loops
#https://www.r-bloggers.com/how-to-write-the-first-for-loop-in-r/

#More about Rbind
#  https://stat.ethz.ch/R-manual/R-devel/library/base/html/cbind.html
```

## Create a StudentSemester variable and decide whether a count is a reasonable thing to do

```{r}
#Create a new variable for Student Semester
MLC$StudentSemester = with(MLC, paste(study_id, semester))

#How many people attended math tutoring more than once within a semester?
  
plot(density(table(MLC$StudentSemester)))
      #there is a lot of variability
      #so it is probably useful to use "count by semester" as a variable
```
## Create an MLC Count Variable - By Semester
  Here, I create an MLC count variable: how many times within a semester did each student attend tutoring?
  For this, each row is a separate observation and thus a separate tutoring attendance event
```{r}
MLC <- MLC %>%
  group_by(StudentSemester) %>% #Unique ID per student semester
  mutate(mlcCount = n())
  #that creates a new variable with 60 unique values

questionr::freq(MLC$mlcCount)
```

## Create an MLC Sum Minutes Variable - By Semester
  Here, I create an MLC sum minutes variable: how many MINUTES within a semester did each student attend tutoring?
  For this, each row is a separate observation and thus a separate tutoring attendance event
```{r}
#Note: Some people are marked "no swipe out" - so this "minutes" variable is compromised and is not a complete representation of how many minutes people were *really* at the math tutoring center

#First, rename the time variable to make it easier to work with
MLC <- MLC %>% 
  rename( mlc_duration = `Time (minutes)`)

#The mlc_duration variable currently has values of "No swipe out" in addition to numeric
  # Convert the variable to numeric
  #This removes the rows that contain text, and replaces them with missing
MLC$mlc_duration <- as.numeric(MLC$mlc_duration)

questionr::freq(MLC$mlc_duration)
  #currently 4 0s

#how many NAs are there?
# MLC %>% 
#   ungroup() %>% 
#   count(is.na(MLC$mlc_duration))
  #599 NAs

  #Recode new mlc_duration variable so that NA is now 0, so that the sum function will work correctly in the next variable creation step 
    #Doing this because the first time I created the sum, it seemed to work
        #but for some reason not everyone was assigned a duration. 
        #So - checking about whether that happened because of NAs?
MLC <- tidyr::replace_na(MLC, list(mlc_duration = 0))

questionr::freq(MLC$mlc_duration)
  #Checking the freq of new 0s. There are 603 0s in the mlc_duration column, where previously there were 599 NAs and 4 0s. Yes, this works!

#Further reading:
  #replace missing "duration" with 0:
# https://stackoverflow.com/questions/19379081/how-to-replace-na-values-in-a-table-for-selected-columns-data-frame-data-tab 

##Now create the minute count variable

MLC <- MLC %>%
  group_by(StudentSemester) %>% #Unique ID per student semester
  mutate(mlc_semester_minutes = sum(mlc_duration))

  #created a new variable with 
      ######### CHECK 
      ##############1594 
      ######### CHECK
      # unique values and no NA

questionr::freq(MLC$mlc_semester_minutes)
```

#Merge SummerA and SummerB into one object
The point of this step is to make it parallel in data structure to my other dataset, which has all Summer variables aggregated together into just "summer"

```{r}
#skipping for now
```

#Transform Long to Wide

```{r}
#First, create a shortened dataset that only includes the variables that I need

short_MLC <- MLC %>% 
  select(study_id, StudentSemester, semester, mlcCount, mlc_semester_minutes)

#Next, keep only unique values of StudentSemester, drop all others
  #StudentSemester is a unique identifier
unique_MLC <- short_MLC %>% 
  dplyr::distinct(StudentSemester, .keep_all = TRUE)

#Spread the unique dataset to create variables for EACH SEMESTER MLC count

mlc_counts<- spread(unique_MLC %>% 
                    group_by(study_id) %>% 
                      select(study_id, semester, mlcCount), semester, mlcCount) 

#Not running the minutes bc decided to eliminate from main analyses
# mlc_times <- spread(unique_MLC %>% 
#                       group_by(study_id) %>% 
#                       select(study_id, semester, mlc_semester_minutes), semester, mlc_semester_minutes)

#Rename the variables so that they make sense in the merged dataset

mlc_counts <- mlc_counts %>% 
            rename(
              mlc_visits_F15 = FS15,
              mlc_visits_S16 = SS16,
              mlc_visits_US16A = US16A,
              mlc_visits_US16B = US16B,
              mlc_visits_FS16 = FS16,
              mlc_visits_SS17 = SS17,
              mlc_visits_US17A = US17A,
              mlc_visits_US17B = US17B,
              mlc_visits_FS17 = FS17,
              mlc_visits_SS18 = SS18
)


#I'm not going to keep these for the Minutes because I decided after writing this code to eliminate the Minutes from my main analyses.
# mlc_times <- mlc_times %>% 
#             rename(
#               mlc_minutes_F15 = FS15,
#               mlc_minutes_S16 = SS16,
#               mlc_minutes_US16A = US16A,
#               mlc_minutes_US16B = US16B,
#               mlc_minutes_FS16 = FS16,
#               mlc_minutes_SS17 = SS17,
#               mlc_minutes_US17A = US17A,
#               mlc_minutes_US17B = US17B,
#               mlc_minutes_FS17 = FS17,
#               mlc_minutes_SS18 = SS18
# )

#join the counts and the times together to be one MLC final dataset

# mlc_final <- left_join(mlc_counts, mlc_times)

```

## Collapse Summer variables into one for each summer
Currently there are 2 variables for summer 2016 (US16A and US16B)
There are also 2 variables for summer 2017 (US17A and US17B)
Here I collapse the summer 2016 variables into 1 and the summer 2017 variables into another variable.

```{r}
mlc_final <- mlc_counts %>% 
            mutate(mlc_visits_US16 = mlc_visits_US16A + mlc_visits_US16B,
                   mlc_visits_US17 = mlc_visits_US17A + mlc_visits_US17B)

#look at the variables by hand to make sure the sums look right - they do. Yay!

#then delete the variables that I don't need
mlc_final <- mlc_final %>%
  select(-c("mlc_visits_US16A", "mlc_visits_US16B", "mlc_visits_US17A", "mlc_visits_US17B"))
    
#now there are 798 obs of 9 variables
```


##Last Save output to a new file

```{r}
#1. Save the result as a CSV

write_csv(mlc_final, "mlc_final.csv")
```