---
title: "Variance Explained July 2019"
author: "Emily Bovee"
date: "7/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This will be the document where I saved all the analyses I did on the CSV files for the variance explained.

For RQ3, I calculated the variance explained right in the documents themselves. However, for these analyses, that won't work because I didn't know how to calculate RMSEs when I was getting the results of those files.

But now I know from the 7.23.19 email from Shahryar that the thing I need to do is to merge the data files together and then get the results into the formula.

These analyses were run on my work computer; the code reflects that accordingly. If needed, I can just change the working directory specification and then run whatever I want.

#Set Working Directory and load files

```{r}
setwd("~/Google Drive/** Dissertation **/VARIANCE EXPLAINED")

library(tidyverse)
library(readr)

#Load Research Question 1 output
model_1a <- read_csv("model_1a.csv")
model_1b <- read_csv("model_1b.csv")
model_1c <- read_csv("model_1c.csv")
model_1d <- read_csv("model_1d.csv")
model_1e <- read_csv("model_1e.csv")

#Load Research Question 2 output
model_2a <- read_csv("model_2a.csv")
model_2b <- read_csv("model_2b.csv")
model_2c <- read_csv("model_2c.csv")
model_2d <- read_csv("model_2d.csv")

```

#Research Question 1: Combining Data & R2

Model 1a
```{r}
#########Model 1a
  #Merge y_hat columns together
pred_1a <- matrix(c(model_1a$yhat_1, model_1a$yhat_2), ncol = 1)

  #merge General y columns togther
obs_1a <- matrix(c(model_1a$y_1, model_1a$y_2), ncol = 1)

  #bind those together

model_1a_R2 <- data.frame(matrix("", nrow = 2088, ncol = 2))
model_1a_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_1a, obs_1a))

#R2
cor(model_1a_R2$yhat, model_1a_R2$y, method = 'pearson')^2

#Here's how I learned how to bind in this way:
#https://www.dummies.com/programming/r/how-to-add-observations-to-a-data-frame-in-r/
```

Model 1b
```{r}
#########Model 1b
  #Merge y_hat columns together
pred_1b <- matrix(c(model_1b$yhat_1, model_1b$yhat_2), ncol = 1)

  #merge General y columns togther
obs_1b <- matrix(c(model_1b$y_1, model_1b$y_2), ncol = 1)

  #bind those together

model_1b_R2 <- data.frame(matrix("", nrow = 2088, ncol = 2))
model_1b_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_1b, obs_1b))

#R2
cor(model_1b_R2$yhat, model_1b_R2$y, method = 'pearson')^2


```


Model 1c
```{r}
#########Model 1c

  #Merge y_hat columns together
pred_1c <- matrix(c(model_1c$yhat_1, model_1c$yhat_2), ncol = 1)

  #merge General y columns togther
obs_1c <- matrix(c(model_1c$y_1, model_1c$y_2), ncol = 1)

  #bind those together

model_1c_R2 <- data.frame(matrix("", nrow = 2088, ncol = 2))
model_1c_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_1c, obs_1c))

#R2
cor(model_1c_R2$yhat, model_1c_R2$y, method = 'pearson')^2


```

Model 1d
```{r}
#########Model 1d
  #Merge y_hat columns together
pred_1d <- matrix(c(model_1d$yhat_1, model_1d$yhat_2), ncol = 1)

  #merge General y columns togther
obs_1d <- matrix(c(model_1d$y_1, model_1d$y_2), ncol = 1)

  #bind those together

model_1d_R2 <- data.frame(matrix("", nrow = 2088, ncol = 2))
model_1d_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_1d, obs_1d))

#R2
cor(model_1d_R2$yhat, model_1d_R2$y, method = 'pearson')^2


```

Model 1e
```{r}
#########Model 1e
  #Merge y_hat columns together
pred_1e <- matrix(c(model_1e$yhat_1, model_1e$yhat_2), ncol = 1)

  #merge General y columns togther
obs_1e <- matrix(c(model_1e$y_1, model_1e$y_2), ncol = 1)

  #bind those together

model_1e_R2 <- data.frame(matrix("", nrow = 2088, ncol = 2))
model_1e_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_1e, obs_1e))

#R2
cor(model_1e_R2$yhat, model_1e_R2$y, method = 'pearson')^2


```
############################
#Research Question 2



Model 2a
```{r}
#########Model 2a
  #Merge y_hat columns together
pred_2a <- matrix(c(model_2a$yhat_1, model_2a$yhat_2, model_2a$yhat_3, model_2a$yhat_4, model_2a$yhat_5), ncol = 1)

  #merge General y columns togther
obs_2a <- matrix(c(model_2a$y_1, model_2a$y_2, model_2a$y_3, model_2a$y_4, model_2a$y_5), ncol = 1)

  #bind those together

model_2a_R2 <- data.frame(matrix("", nrow = 5220, ncol = 2))
model_2a_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_2a, obs_2a))

#R2
cor(model_2a_R2$yhat, model_2a_R2$y, method = 'pearson')^2


```


Model 2b
```{r}
#########Model 2b
  #Merge y_hat columns together
pred_2b <- matrix(c(model_2b$yhat_1, model_2b$yhat_2, model_2b$yhat_3, model_2b$yhat_4, model_2b$yhat_5), ncol = 1)

  #merge General y columns togther
obs_2b <- matrix(c(model_2b$y_1, model_2b$y_2, model_2b$y_3, model_2b$y_4, model_2b$y_5), ncol = 1)

  #bind those together

model_2b_R2 <- data.frame(matrix("", nrow = 5220, ncol = 2))
model_2b_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_2b, obs_2b))

#R2
cor(model_2b_R2$yhat, model_2b_R2$y, method = 'pearson')^2


```


Model 2c
```{r}
#########Model 2c
  #Merge y_hat columns together
pred_2c <- matrix(c(model_2c$yhat_1, model_2c$yhat_2, model_2c$yhat_3, model_2c$yhat_4, model_2c$yhat_5), ncol = 1)

  #merge General y columns togther
obs_2c <- matrix(c(model_2c$y_1, model_2c$y_2, model_2c$y_3, model_2c$y_4, model_2c$y_5), ncol = 1)

  #bind those together

model_2c_R2 <- data.frame(matrix("", nrow = 5220, ncol = 2))
model_2c_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_2c, obs_2c))

#R2
cor(model_2c_R2$yhat, model_2c_R2$y, method = 'pearson')^2

# #variable 1
# cor(model_2c$yhat_1, model_2c$y_1, method = 'pearson')^2
# 
# cor(model_2c$yhat_2, model_2c$y_2, method = 'pearson')^2
# 
# cor(model_2c$yhat_3, model_2c$y_3, method = 'pearson')^2
# 
# cor(model_2c$yhat_4, model_2c$y_4, method = 'pearson')^2
# 
# cor(model_2c$yhat_5, model_2c$y_5, method = 'pearson')^2

```



Model 2d
```{r}
#########Model 2d
  #Merge y_hat columns together
pred_2d <- matrix(c(model_2d$yhat_1, model_2d$yhat_2, model_2d$yhat_3, model_2d$yhat_4, model_2d$yhat_5), ncol = 1)

  #merge General y columns togther
obs_2d <- matrix(c(model_2d$y_1, model_2d$y_2, model_2d$y_3, model_2d$y_4, model_2d$y_5), ncol = 1)

  #bind those together

model_2d_R2 <- data.frame(matrix("", nrow = 5220, ncol = 2))
model_2d_R2[c("yhat", "y")] <- as.data.frame(cbind(pred_2d, obs_2d))

#R2
cor(model_2d_R2$yhat, model_2d_R2$y, method = 'pearson')^2


```
